|

### 2021-07-30 prova_knit_finale.Rmd
added intro

\chapter*{Introduction}

\chapter{Markov Processes and State-Space Models}

@@

\chapter{Linear-Gaussian case}

@@

\chapter{Particle Filter}

@@

\chapter{Applications to Stochastic Volatility Models}

>>>>

\chapter*{Introduction}
State-Space models represent an extremely interesting class of models and they are used extensively, for instance, for inference and forecasting in time-series and macroeconometric analysis. In fact, they allow to model observational data as function of underlying states that cannot be observed. For instance, in macroeconomic models GDP is usually observed while the TFP sequence is the hidden state determining output. In this work we describe techniques, such as Kalman and Particle filter, that allows to approximate probability distributions for the states from the observed data. For example, from the sequence of observed GDP, it is possible to "filter" the TFP factor.\
Furthermore, it is possible to impose a Markov structure to State-Space models, assuming that the states are Markov, i.e., the probability of the current state depends on the entire sequence of past states only through its first lag. The combination of State-Space and Markov structures constitutes a powerful modeling assumption for inference and forecasting.\ \
Withing the Monte Carlo (MC) methods, where inference is based on repeated simulations that allow to approximate the target distributions (needed to compute the estimator of interest), Particle Filter (PF) Algorithms represent the most important class of procedures for State-Space-Markov models. PF allows to sample states period-by-period, updating sequentially the filtering distributions (i.e., the distribution of current state conditional on all the sequence of observations - up to the current period one -), incorporating the information on the likelihood of the extracted states given the sequence of observations, by means of re-weighting.\
Moreover, the PF is a very general algorithm, since it can be applied to non-linear models with non-Gaussian errors, while it does not allow for a closed form solution. On the contrary, the Kalman Filter, is a filtering method that can be applied in the case of linear-Gaussian  State-Space-Markov models, that remains analytically tractable.\ 
In this work we will describe both these algorithms, while our main focus remains on Particle Filter.\ \
In chapter \ref{MP_SSM_chapter} we define precisely Markov processes and State-Space Models. In chapter \ref{KF_chapter} we describe the Kalman Filter and in chapter \ref{PF_chapter} we analyze extensively the Particle Filter. Finally, in chapter \ref{application_chapter}, we apply these methods to stochastic volatility models.\
Finally, we find it useful to integrate each section where an algorithm is described and analyzed from a theoretical perspective with an implementation subsection, using as an example a simple Random-Walk State-Space model. In this way, the reader can visualize the codes and the practical meaning of the Kalman and Particle Filter algorithms, immediately after their theoretical description.

\chapter{Markov Processes and State-Space Models}\label{MP_SSM_chapter}

@@

\chapter{Linear-Gaussian case}\label{KF_chapter}

@@

\chapter{Particle Filter}\label{PF_chapter}

@@

\chapter{Applications to Stochastic Volatility Models}\label{application_chapter}
